---
title: ASA Short Course <br> Second-generation p-values (half-day)
output: 
  html_document:
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
    theme: sandstone
---

# Introduction

+ Jeffrey D. Blume, PhD
    + School of Data Science, University of Virginia
+ Megan H. Murray, almost PhD (July 6th defense)
    + Department of Biostatistics, Vanderbilt University


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=TRUE)

## load packages
library(sgpv)
library(ggplot2)
library(dplyr)
library(kableExtra)
library(FDRestimation)

data(leukstats)
```

# Part I

<!-- install.packages("devtools") -->
<!-- devtools::install_github("weltybiostat/sgpv") -->


```{r}

##########################
###### BOMAMI Trial

## OR
sgpvalue(est.lo=0.967, est.hi=7.286, null.lo=0.9, null.hi=1.11)

## RR
sgpvalue(est.lo=0.953, est.hi=4.589, null.lo=0.9, null.hi=1.11)

## RD
sgpvalue(est.lo=0.003, est.hi=0.352, null.lo=-0.1, null.hi=0.1)
sgpvalue(est.lo=0.003, est.hi=0.352, null.lo=-0.05, null.hi=0.05)

## Logistic REgression

##########################
###### BOMAMI Trial

## Primary OR
sgpvalue(est.lo=1.18, est.hi=20.19, null.lo=0.9, null.hi=1.11)
# sgpvalue(est.lo=1.18, est.hi=20.19, null.lo=0.8, null.hi=1.2)

## Covariates
sgpvalue(est.lo=1.09, est.hi=19.28, null.lo=0.9, null.hi=1.11)
sgpvalue(est.lo=0.86, est.hi=38.29, null.lo=0.9, null.hi=1.11)

sgpvalue(est.lo=0.66, est.hi=8.98, null.lo=0.9, null.hi=1.11)
sgpvalue(est.lo=0.36, est.hi=26.92, null.lo=0.9, null.hi=1.11)

sgpvalue(est.lo=0.95, est.hi=1.11, null.lo=0.9, null.hi=1.11)
sgpvalue(est.lo=0.09, est.hi=1.59, null.lo=0.9, null.hi=1.11)

##########################
###### Leukemia Example

data(leukstats)
plotsgpv(est.lo=leukstats$ci.lo, est.hi=leukstats$ci.hi,
		null.lo=-0.3, null.hi=0.3,
		set.order=order(leukstats$p.value),
		x.show=7000,
		plot.axis=c("TRUE","FALSE"),
		null.pt=0, outline.zone=TRUE,
		title.lab="Leukemia Example", y.lab="Fold Change (base 10)",
		x.lab="Classical p-value ranking",
		legend.on=TRUE)
axis(side=2,at=round(log(c(1/1000,1/100,1/10,1/2,1,2,10,100,1000),
	base=10),2),labels=c("1/1000","1/100","1/10","1/2",1,2,10,100,1000),
	las=2)

##########################
###### FDR rates

fdrisk(sgpval = 0,  null.lo = log(1/1.1), null.hi = log(1.1),  std.err = 0.8,  null.weights = 'Uniform',  null.space = c(log(1/1.1), log(1.1)),  alt.weights = 'Uniform',  alt.space = 2 + c(-1,1)*qnorm(1-0.05/2)*0.8,  interval.type = 'confidence',  interval.level = 0.05)
fdrisk(sgpval = 0,  null.lo = log(1/1.1), null.hi = log(1.1),  std.err = 0.8,  null.weights = 'Point',  null.space = 0,  alt.weights = 'TruncNormal',  alt.space = 2 + c(-1,1)*qnorm(1-0.041/2)*0.8,  interval.type = 'likelihood',  interval.level = 1/8)
fdrisk(sgpval = 0,  null.lo = log(1/1.5), null.hi = log(1.5),  std.err = 0.8,  null.weights = 'Point',  null.space = 0,  alt.weights = 'Uniform',  alt.space = 2.5 + c(-1,1)*qnorm(1-0.041/2)*0.8,  interval.type = 'likelihood',  interval.level = 1/8)
fdrisk(sgpval = 1,  null.lo = log(1/1.5), null.hi = log(1.5),  std.err = 0.15,  null.weights = 'Uniform',  null.space = 0.01 + c(-1,1)*qnorm(1-0.041/2)*0.15,  alt.weights = 'Uniform',  alt.space = c(log(1.5), 1.25*log(1.5)),  interval.type = 'likelihood',  interval.level = 1/8)

###
##
#
```

# Part II

## FDRestimation

Download the package from:
https://cran.r-project.org/package=FDRestimation

OR https://github.com/murraymegan/FDRestimation 

Our corresponding paper: https://arxiv.org/abs/2010.04680 

PDF: https://arxiv.org/pdf/2010.04680.pdf 


```{r}
p.fdr.output = p.fdr(leukstats$p.value)

head(p.fdr.output$`Results Matrix`)

summary(p.fdr.output)

plot(p.fdr.output)

get.pi0(leukstats$p.value, estim.method = "last.hist")
get.pi0(leukstats$p.value, estim.method = "storey")
get.pi0(leukstats$p.value, estim.method = "set.pi0", set.pi0=0.8)

plot(p.fdr.output, xlim=c(1100,1600), ylim=c(0, 0.2))

which(-1*(p.fdr.output$`Results Matrix`$`Adjusted p-values`)+p.fdr.output$fdrs>0.001)

#Benjamini-Yeukatelli
p.fdr.output = p.fdr(leukstats$p.value, adjust.method = "BY")

head(p.fdr.output$`Results Matrix`)

plot(p.fdr.output)

plot(p.fdr.output, xlim=c(2000,2400), legend.on = FALSE)
```

### Simple Example

```{r}
pvalues=c(0.005,0.049,0.05,0.051,0.7)
zvalues=qnorm(pvalues/2, lower.tail = FALSE)

p.fdr.output = p.fdr(pvalues)

adj.pvalues= p.fdr.output$`Results Matrix`$`Adjusted p-values`
adj.fdrs= p.fdr.output$fdrs

single.fdr = c(p.fdr(pvalues=pvalues[1],zvalue=zvalues[1]),
               p.fdr(pvalues=pvalues[2],zvalue=zvalues[2]),
               p.fdr(pvalues=pvalues[3],zvalue=zvalues[3]),
               p.fdr(pvalues=pvalues[4],zvalue=zvalues[4]),
               p.fdr(pvalues=pvalues[5],zvalue=zvalues[5]))
```

```{r}
df = data.frame("Raw p-values"= pvalues, 
                "Z-values" = zvalues,
                "BH Adj p-values" = adj.pvalues,
                "BH FDRs" = adj.fdrs,
                "Single lower bound FDRs" = single.fdr)

colnames(df) = c("Raw p-values", 
                "Z-values",
                "BH Adj p-values"  ,
                "BH FDRs" ,
                "Single lower bound FDRs" )
                
kable(round(df,3))%>%
  kable_styling(c("striped", "bordered"))
```

### Compare FDRestimation to p.adjust

```{r}
set.seed(999)
pi0 <- 0.8
pi1 <- 1-pi0
n <- 10
n.0 <- ceiling(n*pi0)
n.1 <- n-n.0

sim.data <- c(rnorm(n.1,2,1),rnorm(n.0,0,1))
sim.data.p <- 2*pnorm(-abs(sim.data))

p.adjust.output = p.adjust(sim.data.p, method="fdr")

fdr.output = p.fdr(pvalues=sim.data.p, adjust.method="BH")

head(data.frame("Raw p-values"= sim.data.p,"p.adjust FDRs"=p.adjust.output,"p.fdr adj p-values"=fdr.output$`Results Matrix`$`Adjusted p-values`,"p.fdr FDRs"=fdr.output$fdrs))

plot(rank(sim.data.p, ties="random"), 
     fdr.output$`Results Matrix`$`Adjusted p-values`, 
     pch=17, 
     col="dodgerblue", 
     cex=2, 
     ylim=c(0,1),
     main="Comparison Plot", 
     xlab="Rank of p-values",
     ylab="")
points(rank(sim.data.p, ties="random"),p.adjust.output, pch=20, col="pink")
points(rank(sim.data.p, ties="random"),fdr.output$fdrs, pch=20, col="firebrick")
legend("bottomright", c("p.adjust FDRs", "p.fdr FDRs","p.fdr Adjusted p-values"), col=c("pink", "firebrick","dodgerblue"), pch=c(20,20,17))
```

### Different Null proportion

```{r}
set.seed(999)
pi0 <- 0.8
pi1 <- 1-pi0
n <- 100
n.0 <- ceiling(n*pi0)
n.1 <- n-n.0

sim.data <- c(rnorm(n.1,2,1),rnorm(n.0,0,1))
sim.data.p <- 2*pnorm(-abs(sim.data))

fdr.output = p.fdr(pvalues=sim.data.p, adjust.method="BH")

plot(fdr.output)

fdr.output = p.fdr(pvalues=sim.data.p, adjust.method="BH", set.pi0=0.8)

plot(fdr.output)
```

```{r}
get.pi0(sim.data.p, estim.method = "set.pi0", set.pi0=0.7)
get.pi0(sim.data.p, estim.method = "last.hist")
get.pi0(sim.data.p, estim.method = "storey")

pi0 <- 0.8
pi1 <- 1-pi0
n <- 10000
n.0 <- ceiling(n*pi0)
n.1 <- n-n.0

sim.data <- c(rnorm(n.1,2,1),rnorm(n.0,0,1))
sim.data.p <- 2*pnorm(-abs(sim.data))

get.pi0(sim.data.p, estim.method = "set.pi0", set.pi0=0.7)
get.pi0(sim.data.p, estim.method = "last.hist")
get.pi0(sim.data.p, estim.method = "storey")
```

#### Internal Code: Last Histogram Height Method

```{r}
try.hist = hist(sim.data.p, breaks="scott")
try.mids = try.hist$mids
try.count = try.hist$counts

if(tail(try.mids,1)<0.5){
  pi0.hat=0
}else{
  pi0.hat = min(tail(try.count,1)*length(try.mids)/sum(try.count),1)
}

pi0.hat
```
